\LARGE \bf Mean Estimation from One-Bit Measurements

### 2022-05-04 proof-sgd.tex
arranged appendix in two-columns

Consider the following simplified version of
\cite[Thm. 4]{polyak1992acceleration}:
\begin{corollary}{\cite[Thms. 3 \& 4]{polyak1992acceleration}}
  \label{corollary:polyak-juditsky}
  Let $\varphi : \R \to \R$ and $\{Z_i\}$ be i.i.d.\ \newtext{zero-mean} random
  variables, and
  \begin{equation*}
  X_i = \theta + Z_i.
  \end{equation*}
  Define
  \begin{align}
    \begin{split}
      \theta_i & = \theta_{i-1} + \gamma_i \varphi(X_i - \theta_{i-1}), \\
      \bar{\theta}_n & = \frac{1}{n} \sum_{i=0}^{n-1} \theta_i, 
    \end{split}
    \label{eq:Polyak_Juditsky_alg}
  \end{align}
  where in addition,
  \begin{enumerate}[(i)]
  \item There exists $K_1$ such that $\left| \varphi(x) \right| \leq
    K_1(1+|x|)$ for all $x\in \R$.
  \item The sequence $\left\{ \gamma_i \right\}_{i=1}^\infty$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
  \item \label{item:zero-gradient}
    The function $\psi(x) \defeq \ex{ \varphi(x+Z_1)}$
    satisfies $\psi(0) = 0$ and
    $x\psi(x) > 0$ for $x\neq 0$. Moreover, $\psi$ is differentiable
    at 0 with $\psi'(0) > 0$ and there exists
    $K_2$, $0 < \lambda \leq 1$, \newtext{and $r>0$}, such that
    \begin{equation}
      \label{eqn:local-hessian-psi}
        \left| \psi(x) - \psi'(0)x \right|\leq K_2 |x|^{1+\lambda}
    \end{equation}
    \newtext{for all $|x|<r$. }
  \item The function 
    $\chi(x) \defeq \ex{\varphi^2(x+Z_1)}$ is continuous at zero. 
  \end{enumerate}
  Then $\bar{\theta}_n \cas \theta$ and $ \sqrt{n}({\theta}_n - \theta)
  \cd \normal(0,V)$ for
  $V = \frac{ \chi(0)} {\psi'(0)^2}$.
\end{corollary}

Using the notation in Corollary~\ref{corollary:polyak-juditsky}, we set
$\varphi(x) = \sgn(x)$ and $Z_i = X_i - \theta$, where $\theta =
\mbox{med}(P)$. Without loss of generality and for notational convenience,
we assume for the remainder of this derivation that $\theta = 0$.
As a consequence, we have $\mbox{med}(Z) = 0$,
and $\chi(x) = \ex{ \sgn^2(x+Z_1) }= 1$, so
$\chi(0) = 1$. In addition,
\begin{align*}
  \psi(x) & = \ex{ \sgn(x+ Z_1) } =
  P(Z \ge -x) - P(Z < -x) = 1 - 2 P(Z \le -x).
  %%   \int_{-\infty}^\infty \sgn(x+z) f(z) dz \\
  %% & = \int_{-x}^\infty f(z) dz -\int_{-\infty}^{-x} f(z) dz. 
\end{align*}
Using that $P$ has a density $f$ near its median, it follows that $\psi'(x)
= 2f(-x)$ and thus $\psi'(0) = 2f(0) > 0$.  We may now verify that the
conditions in Corollary~\ref{corollary:polyak-juditsky} hold for $\lambda =
1$. Condition~(i) is obvious, and the convexity of $|\cdot|$ gives most of
condition~(iii) excepting inequality~\eqref{eqn:local-hessian-psi}. For
that, note that as $f$ is Lipschitz near 0 with constant $\lip_0(f)$, we
have for small $x$ that
% AK: Here I changed \pm to \leq and \geq
\begin{equation*}
  \psi(x) = 2 \int_0^x f(-t) dt
  \leq 2 \int_0^x \left[f(0) + \lip_0(f) t \right] dt
  = 2 f(0) x + \lip_0(f) x^2
  = \psi'(0) x + \lip_0(f) x^2,
\end{equation*}
\begin{equation*}
  \psi(x) = 2 \int_0^x f(-t) dt
  \geq 2 \int_0^x \left[f(0) - \lip_0(f) t \right] dt
  = 2 f(0) x - \lip_0(f) x^2
  = \psi'(0) x - \lip_0(f) x^2,
\end{equation*}
so that condition~(iii) holds.
As evidently $\chi(0) / \psi'(0)^2 = \frac{1}{4 f(0)^2}$,
Corollary~\ref{corollary:polyak-juditsky} gives
Theorem~\ref{thm:sgd}\eqref{item:normal-sgd}.

@@

These three definitions are linked in our case by a few important results.
First~\cite[Theorem 7.2]{VanDerVaart98}, if $\{P_\theta\}$ is QMD
(Def.~\ref{definition:qmd}) at the point $\theta$, then it is locally
asymptotically normal with $Z_n = \frac{1}{\sqrt{n}} \sum_{i = 1}^n
\score_\theta(X_i)$ and information matrix $I_\theta =
\E_\theta[\score_\theta \score_\theta^\top]$. Moreover, in any family
$\{P_\theta\}$ that is LAN (Def.~\ref{definition:lan}) at $\theta$, if $T_n$
is a regular estimator (Def.~\ref{definition:regular-estimator}) at $\theta$
with limiting distribution $Q$, then for any bounded, symmetric,
quasi-convex loss $L$ and $c < \infty$,
\begin{equation}
  \label{eqn:limit-law-regular}
  \limsup_n \sup_{\norm{h} \le c}
  \E_{P_{\theta + h/\sqrt{n}}}
  \left[L(\sqrt{n}(T_n - \theta - h / \sqrt{n}))\right]
  = \E[L(W)] ~~ \mbox{for~}W \sim Q
\end{equation}
(see Beran~\cite{beran1995role}, Eq.~(4.2)).
%
Thus, we show two results: first,
that the family $\{P_\theta\}$ of distributions
defined by the shifted densities $\{f(\cdot - \theta)\}_{\theta \in \R}$
is quadratic-mean-differentiable at any $\theta$,
and second, that $\bar{\theta}_n$ is regular and
asymptotically normal.
The combination evidently gives the theorem.

@@

  We now verify that the setting of Theorem~\ref{thm:sgd}
  (and Lemma~\ref{lemma:sgd-median-regular}) satisfies the
  conditions of Corollary~\ref{corollary:normal-expansion}. First, we have
  the obvious fact that
  \begin{equation*}
    |\sgn(z) - \sgn(x + z)|
    \le 2 \cdot \indic{|x| \ge |z|}.
  \end{equation*}
  Recalling that the density $f$ is Lipschitz with median 0,
  for $\varphi(z) = \sgn(z)$, and $Z = X - \theta$ distributed with
  density $f$, we have
  \begin{align*}
    \ex{ \left| \varphi(Z) - \varphi(x + Z) \right| }
    \leq 2 \Prob\left( |Z_1| \le |x|  \right)
    = 2 \int_{-|x|}^{|x|} f(t) dt
    \le 4 f(0)|x| + 2 \int_{-|x|}^{|x|}
    \lip(f) t dt
    = 4 f(0) |x| + 2 \lip(f) x^2
  \end{align*}
  where $\lip(f)$ is the Lipschitz constant of $f$.
\newtext{It follows that condition~\eqref{eqn:additional-local-smooth} holds.}  
   %For large $x$ we always have  $\E[|\varphi(Z) - \varphi(x + Z)|] \le 2$, so that condition~\eqref{eqn:additional-local-smooth} holds.
  In addition, we have
  \begin{align*}
\newtext{ \E_p[\varphi(Z_1) \score_0(Z_1)] } & = 
    \int_{\R} \varphi(x) f'(x ) dx
    = \int_{\R} \sgn(x) f'(x ) dx\\
     &  =
    \int_0^\infty f'(x ) dx
    - \int_{-\infty}^0 f'(x ) dx
    = -2 f(0) = -\psi'(0).
  \end{align*}
  Corollary~\ref{corollary:normal-expansion}
  now implies the convergence \eqref{eqn:sgd-median-regular}.
\end{proof}

@@

\begin{lemma}
  The expansion~\eqref{eq:normal_expansion_lem} holds for the process
  $\bar{\Delta}^1_n$ defined by the iteration
  \begin{align}
    \label{eqn:polyak-expansion_lem1_alg}
    & \Delta_i^1  = \Delta_{i-1}^1 - \gamma_i \psi'(0) \Delta_{i-1}^1 - \gamma_i \varphi(Z_i), \qquad
    \Delta_0^1 = \Delta_0\\
    & \bar{\Delta}^1_n = \frac{1}{n}\sum_{i=0}^{n-1} \Delta^1_i.
    \nonumber
  \end{align}
\end{lemma}
\begin{proof}
  To prove this claim, use Lemma~\ref{lemma:polyak-expansion} with $A =
  \psi'(0)$ and $\xi_i = -\varphi(Z_i)$, which by
  condition~\eqref{item:zero-gradient} in
  Corollary~\ref{corollary:polyak-juditsky} gives that
  $\E[\xi_i] = 0$ and that the $\xi_i$ are independent.
  The first term $\alpha_n \Delta_0^1 / \gamma_0 \sqrt{n} \to 0$
  in Eq.~\eqref{eqn:polyak-expansion}. In addition,
  by independence and that the $\xi_i$ are mean-zero, we have
  \begin{align*}
    \ex{ \left( \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} w_i^n \xi_i \right)^2 }
    & = \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \ex{ \xi_i^2} + \frac{1}{n}  \sum_{i\neq j}^n w_i^n w_j^n \ex{ \xi_i \xi_j} \\
    & = \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \ex{ \varphi(Z_i)^2} = \chi(0) \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \to 0
  \end{align*}
  by Lemma~\ref{lemma:polyak-expansion}.
  Thus, the expansion~\eqref{eqn:polyak-expansion} in
  Lemma~\ref{lemma:polyak-expansion} gives
  \begin{equation*}
    \sqrt{n} \bar{\Delta}^1_n
    = -\frac{1}{\sqrt{n}} \frac{1}{\psi'(0)}
    \sum_{i=1}^{n-1} \varphi(Z_i)+ o_{P,n}(1)
  \end{equation*}
  as desired.
\end{proof}

We then have the following asymptotic equivalence.
\begin{lemma}
  The sequences $\bar{\Delta}_n$ and $\bar{\Delta}^1_n$ are asymptotically
  equivalent, meaning that
  $\sqrt{n} (\bar{\Delta}_n - \bar{\Delta}_n^1) \cp 0$.
\end{lemma}
\begin{proof}
  From the recursions~\eqref{eq:Polyak_Juditsky_alg} and
  \eqref{eqn:polyak-expansion_lem1_alg}, the difference $\delta_i = \Delta_i
  - \Delta_i^1$ satisfies
  \begin{equation*}
  \delta_i = \delta_{i-1} - \gamma_i \psi'(0) \delta_{i-1} + \gamma_i \left( \psi'(0) \Delta_{i-1}  + \varphi(Z_i) - \varphi(\Delta_{i-1} + Z_i) \right),
  \end{equation*}
  where $\delta_0 = 0$. Applying Lemma~\ref{lemma:polyak-expansion} with the
  choices $\xi_i = \psi'(0) \Delta_{i-1} + \varphi(Z_i) -
  \varphi(\Delta_{i-1} + Z_i)$ yields
  \begin{align}
    \sqrt{n}\bar{\delta}_n
    & = \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1}
    \left( \frac{1}{\psi'(0)} + w_i^n \right)  \xi_i  \nonumber \\
    & = \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1}
    \left( \frac{1}{\psi'(0)} + w_i^n \right)
    \left( \psi'(0) \Delta_{i-1}  - \psi(\Delta_{i-1}) \right)
    \label{eq:PJ_proof1} \\
    & \qquad ~ + 
    \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} \left( \frac{1}{\psi'(0)}
    + w_i^n \right)
    \left( \psi(\Delta_{i-1})  + \varphi(Z_i) - \varphi(\Delta_{i-1}+Z_i)
    \right) \label{eq:PJ_proof2}
  \end{align}
  For the term \eqref{eq:PJ_proof1},
  the assumption~\eqref{eqn:local-hessian-psi} that
  $|\psi(x) - \psi'(0) x| = O(x^{1 + \lambda})$
  and that $\sup_{i,n} |w_i^n| < \infty$ by Lemma~\ref{lemma:polyak-expansion}
  give that there exists $K < \infty$ such that
  $|\psi'(0)^{-1} + w_i^n| |\psi'(0) \Delta_{i-1} - \psi(\Delta_{i-1})|
  \le K |\Delta_i|^{1 + \lambda}$.
  Lemma~\ref{lemma:converging-power-sum} gives that
  $\sum_{i = 1}^n
  \frac{1}{\sqrt{i}} |\Delta_i|^{1 + \lambda} < \infty$,
  and so the Kronecker lemma gives that
  \begin{equation*}
    \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} \left( \frac{1}{\psi'(0)}  + w_i^n \right)  \left( \psi'(0) \Delta_{i-1}  - \psi(\Delta_{i-1}) \right) \cas 0.
  \end{equation*}

>>>>

Consider the following simplified version of
\cite[Thm. 4]{polyak1992acceleration}:
\begin{corollary}{\cite[Thms. 3 \& 4]{polyak1992acceleration}}
  \label{corollary:polyak-juditsky}
  Let $\varphi : \R \to \R$ and $\{Z_i\}$ be i.i.d.\ zero-mean random
  variables, and
  \begin{equation*}
  X_i = \theta + Z_i.
  \end{equation*}
  Define
  \begin{align}
    \begin{split}
      \theta_i & = \theta_{i-1} + \gamma_i \varphi(X_i - \theta_{i-1}), \\
      \bar{\theta}_n & = \frac{1}{n} \sum_{i=0}^{n-1} \theta_i, 
    \end{split}
    \label{eq:Polyak_Juditsky_alg}
  \end{align}
  where in addition,
  \begin{enumerate}[(i)]
  \item There exists $K_1$ such that $\left| \varphi(x) \right| \leq
    K_1(1+|x|)$ for all $x\in \R$.
  \item The sequence $\left\{ \gamma_i \right\}_{i=1}^\infty$ satisfies
    condition~\eqref{eqn:lazy-gamma}.
  \item \label{item:zero-gradient}
    The function $\psi(x) \defeq \ex{ \varphi(x+Z_1)}$
    satisfies $\psi(0) = 0$ and
    $x\psi(x) > 0$ for $x\neq 0$. Moreover, $\psi$ is differentiable
    at 0 with $\psi'(0) > 0$ and there exists
    $K_2$, $0 < \lambda \leq 1$, \newtext{and $r>0$}, such that
    \begin{equation}
      \label{eqn:local-hessian-psi}
        \left| \psi(x) - \psi'(0)x \right|\leq K_2 |x|^{1+\lambda}
    \end{equation}
    \newtext{for all $|x|<r$. }
  \item The function 
    $\chi(x) \defeq \ex{\varphi^2(x+Z_1)}$ is continuous at zero. 
  \end{enumerate}
  Then $\bar{\theta}_n \cas \theta$ and $ \sqrt{n}({\theta}_n - \theta)
  \cd \normal(0,V)$ for
  $V = \frac{ \chi(0)} {\psi'(0)^2}$.
\end{corollary}

Using the notation in Corollary~\ref{corollary:polyak-juditsky}, we set
$\varphi(x) = \sgn(x)$ and $Z_i = X_i - \theta$, where $\theta =
\mbox{med}(P)$. Without loss of generality and for notational convenience,
we assume for the remainder of this derivation that $\theta = 0$.
As a consequence, we have $\mbox{med}(Z) = 0$,
and $\chi(x) = \ex{ \sgn^2(x+Z_1) }= 1$, so
$\chi(0) = 1$. In addition,
\begin{align*}
  \psi(x) & = \ex{ \sgn(x+ Z_1) } =
  P(Z \ge -x) - P(Z < -x) = 1 - 2 P(Z \le -x).
  %%   \int_{-\infty}^\infty \sgn(x+z) f(z) dz \\
  %% & = \int_{-x}^\infty f(z) dz -\int_{-\infty}^{-x} f(z) dz. 
\end{align*}
Using that $P$ has a density $f$ near its median, it follows that $\psi'(x)
= 2f(-x)$ and thus $\psi'(0) = 2f(0) > 0$.  We may now verify that the
conditions in Corollary~\ref{corollary:polyak-juditsky} hold for $\lambda =
1$. Condition~(i) is obvious, and the convexity of $|\cdot|$ gives most of
condition~(iii) excepting inequality~\eqref{eqn:local-hessian-psi}. For
that, note that as $f$ is Lipschitz near $0$ with constant $\lip_0(f)$, we 
have for small $x$ that
\begin{align*}
  \psi(x) & = 2 \int_0^x f(-t) dt
  \leq 2 \int_0^x \left[f(0) + \lip_0(f) t \right] dt \\
  & = 2 f(0) x + \lip_0(f) x^2 = \psi'(0) x + \lip_0(f) x^2,
\end{align*}
\begin{align*}
  \psi(x) & = 2 \int_0^x f(-t) dt
  \geq 2 \int_0^x \left[f(0) - \lip_0(f) t \right] dt
  \\
  & = 2 f(0) x - \lip_0(f) x^2 = \psi'(0) x - \lip_0(f) x^2,
\end{align*}
so that condition~(iii) holds.
As evidently $\chi(0) / \psi'(0)^2 = \frac{1}{4 f(0)^2}$,
Corollary~\ref{corollary:polyak-juditsky} gives
Theorem~\ref{thm:sgd}\eqref{item:normal-sgd}.

@@

These three definitions are linked in our case by a few important results.
First~\cite[Theorem 7.2]{VanDerVaart98}, if $\{P_\theta\}$ is QMD
(Def.~\ref{definition:qmd}) at the point $\theta$, then it is locally
asymptotically normal with $Z_n = \frac{1}{\sqrt{n}} \sum_{i = 1}^n
\score_\theta(X_i)$ and information matrix $I_\theta =
\E_\theta[\score_\theta \score_\theta^\top]$. Moreover, in any family
$\{P_\theta\}$ that is LAN (Def.~\ref{definition:lan}) at $\theta$, if $T_n$
is a regular estimator (Def.~\ref{definition:regular-estimator}) at $\theta$
with limiting distribution $Q$, then for any bounded, symmetric,
quasi-convex loss $L$ and $c < \infty$,
\begin{align}
  \limsup_n \sup_{\norm{h} \le c}
  & \E_{P_{\theta + h/\sqrt{n}}}
  \left[L(\sqrt{n}(T_n - \theta - h / \sqrt{n}))\right] \nonumber \\
  & = \E[L(W)] ~~ \mbox{for~}W \sim Q \label{eqn:limit-law-regular}
\end{align}
(see Beran~\cite{beran1995role}, Eq.~(4.2)).
%
Thus, we show two results: first,
that the family $\{P_\theta\}$ of distributions
defined by the shifted densities $\{f(\cdot - \theta)\}_{\theta \in \R}$
is quadratic-mean-differentiable at any $\theta$,
and second, that $\bar{\theta}_n$ is regular and
asymptotically normal.
The combination evidently gives the theorem.

@@

  We now verify that the setting of Theorem~\ref{thm:sgd}
  (and Lemma~\ref{lemma:sgd-median-regular}) satisfies the
  conditions of Corollary~\ref{corollary:normal-expansion}. First, we have
  the obvious fact that
  \begin{equation*}
    |\sgn(z) - \sgn(x + z)|
    \le 2 \cdot \indic{|x| \ge |z|}.
  \end{equation*}
  Recalling that the density $f$ is Lipschitz with median 0,
  for $\varphi(z) = \sgn(z)$, and $Z = X - \theta$ distributed with
  density $f$, we have
  \begin{align*}
    & \ex{ \left| \varphi(Z) - \varphi(x + Z) \right| }
    \leq 2 \Prob\left( |Z_1| \le |x|  \right) \\
    & \quad = 2 \int_{-|x|}^{|x|} f(t) dt
    \le 4 f(0)|x| + 2 \int_{-|x|}^{|x|}
    \lip(f) t dt \\
     & \quad = 4 f(0) |x| + 2 \lip(f) x^2
  \end{align*}
  where $\lip(f)$ is the Lipschitz constant of $f$.
\newtext{It follows that condition~\eqref{eqn:additional-local-smooth} holds.}  
   %For large $x$ we always have  $\E[|\varphi(Z) - \varphi(x + Z)|] \le 2$, so that condition~\eqref{eqn:additional-local-smooth} holds.
  In addition, we have
  \begin{align*}
& \newtext{ \E_p[\varphi(Z_1) \score_0(Z_1)] } = 
    \int_{\R} \varphi(x) f'(x ) dx
    = \int_{\R} \sgn(x) f'(x ) dx\\
     & \quad  =
    \int_0^\infty f'(x ) dx
    - \int_{-\infty}^0 f'(x ) dx
    = -2 f(0) = -\psi'(0).
  \end{align*}
  Corollary~\ref{corollary:normal-expansion}
  now implies the convergence \eqref{eqn:sgd-median-regular}.
\end{proof}

@@

\begin{lemma}
  The expansion~\eqref{eq:normal_expansion_lem} holds for the process
  $\bar{\Delta}^1_n$ defined by the iteration
  \begin{align}
    & \Delta_i^1  = \Delta_{i-1}^1 - \gamma_i \psi'(0) \Delta_{i-1}^1 - \gamma_i \varphi(Z_i), \qquad
    \Delta_0^1 = \Delta_0 \nonumber\\
    & \bar{\Delta}^1_n = \frac{1}{n}\sum_{i=0}^{n-1} \Delta^1_i.
    \label{eqn:polyak-expansion_lem1_alg}
  \end{align}
\end{lemma}
\begin{proof}
  To prove this claim, use Lemma~\ref{lemma:polyak-expansion} with $A =
  \psi'(0)$ and $\xi_i = -\varphi(Z_i)$, which by
  condition~\eqref{item:zero-gradient} in
  Corollary~\ref{corollary:polyak-juditsky} gives that
  $\E[\xi_i] = 0$ and that the $\xi_i$ are independent.
  The first term $\alpha_n \Delta_0^1 / \gamma_0 \sqrt{n} \to 0$
  in Eq.~\eqref{eqn:polyak-expansion}. In addition,
  by independence and that the $\xi_i$ are mean-zero, we have
  \begin{align*}
    & \ex{ \left( \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} w_i^n \xi_i \right)^2 } \\
     & \quad = \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \ex{ \xi_i^2} + \frac{1}{n}  \sum_{i\neq j}^n w_i^n w_j^n \ex{ \xi_i \xi_j} \\
    & \quad = \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \ex{ \varphi(Z_i)^2} = \chi(0) \frac{1}{n}  \sum_{i=1}^n (w_i^n)^2 \to 0
  \end{align*}
  by Lemma~\ref{lemma:polyak-expansion}.
  Thus, the expansion~\eqref{eqn:polyak-expansion} in
  Lemma~\ref{lemma:polyak-expansion} gives
  \begin{equation*}
    \sqrt{n} \bar{\Delta}^1_n
    = -\frac{1}{\sqrt{n}} \frac{1}{\psi'(0)}
    \sum_{i=1}^{n-1} \varphi(Z_i)+ o_{P,n}(1)
  \end{equation*}
  as desired.
\end{proof}

We then have the following asymptotic equivalence.
\begin{lemma}
  The sequences $\bar{\Delta}_n$ and $\bar{\Delta}^1_n$ are asymptotically
  equivalent, meaning that
  $\sqrt{n} (\bar{\Delta}_n - \bar{\Delta}_n^1) \cp 0$.
\end{lemma}
\begin{proof}
  From the recursions~\eqref{eq:Polyak_Juditsky_alg} and
  \eqref{eqn:polyak-expansion_lem1_alg}, the difference $\delta_i = \Delta_i
  - \Delta_i^1$ satisfies
  \begin{align*}
  & \delta_i = \delta_{i-1} - \gamma_i \psi'(0) \delta_{i-1}  \\
  & \qquad + \gamma_i \left( \psi'(0) \Delta_{i-1}  + \varphi(Z_i) - \varphi(\Delta_{i-1} + Z_i) \right),
  \end{align*}
  where $\delta_0 = 0$. Applying Lemma~\ref{lemma:polyak-expansion} with the
  choices $\xi_i = \psi'(0) \Delta_{i-1} + \varphi(Z_i) -
  \varphi(\Delta_{i-1} + Z_i)$ yields
  \begin{align}
    & \sqrt{n}\bar{\delta}_n
    = \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1}
    \left( \frac{1}{\psi'(0)} + w_i^n \right)  \xi_i  \nonumber \\
    & \quad = \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1}
    \left( \frac{1}{\psi'(0)} + w_i^n \right)
    \left( \psi'(0) \Delta_{i-1}  - \psi(\Delta_{i-1}) \right)
    \label{eq:PJ_proof1} \\
    & \qquad ~ + 
    \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} \left( \frac{1}{\psi'(0)}
    + w_i^n \right) \label{eq:PJ_proof2} \\
    & \qquad \qquad \times 
    \left( \psi(\Delta_{i-1})  + \varphi(Z_i) - \varphi(\Delta_{i-1}+Z_i)
    \right) \nonumber 
  \end{align}
  For the term \eqref{eq:PJ_proof1},
  the assumption~\eqref{eqn:local-hessian-psi} that
  $|\psi(x) - \psi'(0) x| = O(x^{1 + \lambda})$
  and that $\sup_{i,n} |w_i^n| < \infty$ by Lemma~\ref{lemma:polyak-expansion}
  give that there exists $K < \infty$ such that
  $|\psi'(0)^{-1} + w_i^n| |\psi'(0) \Delta_{i-1} - \psi(\Delta_{i-1})|
  \le K |\Delta_i|^{1 + \lambda}$.
  Lemma~\ref{lemma:converging-power-sum} gives that
  $\sum_{i = 1}^n
  \frac{1}{\sqrt{i}} |\Delta_i|^{1 + \lambda} < \infty$,
  and so the Kronecker lemma gives that
  \begin{equation*}
    \frac{1}{\sqrt{n}} \sum_{i=1}^{n-1} \left( \frac{1}{\psi'(0)}  + w_i^n \right)  \left( \psi'(0) \Delta_{i-1}  - \psi(\Delta_{i-1}) \right) \cas 0.
  \end{equation*}

  The term \eqref{eq:PJ_proof2} is somewhat more challenging to control.
  We define
  \begin{equation*}
    \epsilon_i \defeq \psi(\Delta_{i-1}) + \varphi(Z_i)
    - \varphi(\Delta_{i-1}+Z_i),
  \end{equation*}
  and let
  $\mc{F}_i = \sigma(Z_1, \ldots, Z_i)$ be the $\sigma$-field of the randomness
  through time $i$. We use a square integrable martingale convergence
  theorem~\cite[Exercise~5.3.35]{Dembo16}. Noting that
  $\Delta_i \in \mc{F}_i$, we have
  \begin{align}
    & \E[\epsilon_i^2 \mid \mc{F}_{i-1}] \nonumber \\
    & \quad = \E[(\psi(\Delta_{i-1}) + \varphi(Z_i)
      - \varphi(\Delta_{i-1} + Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \quad  \le 2 \psi(\Delta_{i-1})^2
    + 2 \E[(\varphi(\Delta_{i-1} + Z_i) - \varphi(Z_i))^2 \mid \mc{F}_{i-1}]
    \nonumber \\
    & \quad \le
    K \left[|\Delta_{i-1}|^{1 + \lambda}
      + |\Delta_{i-1}|^\lambda
      + \Delta_{i-1}^2 \right],
    \label{eqn:bound-psi-error-expectations}
  \end{align}
  where inequality~\eqref{eqn:bound-psi-error-expectations} follows by the
  conditions~\eqref{eqn:local-hessian-psi}
  and~\eqref{eqn:additional-local-smooth}, and $\E[\varepsilon_i \mid
    \mc{F}_{i-1}] = 0$ for all $i$ by definition of $\psi(x) = \E[\varphi(x
    + Z)]$ and that $\psi(0) = 0$.  We now control the expectations of these
  quantities. For $R<\infty$, define the the stopping time $\tau_R \defeq \inf
  \{i : |\Delta_i| > R\}$, which satisfies $\{\tau_R \le i\} \in \mc{F}_i$
  for each $i$. Then using~\cite[Eq.~(A13-A14)]{polyak1992acceleration}, we have
  \begin{equation*}
    \E[\Delta_i^2 \indic{\tau_R > i}] \le K \gamma_i,
  \end{equation*}
  and so inequality~\eqref{eqn:bound-psi-error-expectations} gives that
  \begin{align*}
    & \E\bigg[\sum_{i = 1}^\infty \frac{1}{i}
      |\varepsilon_i|^2 \indic{\tau_R > n} \bigg]
    \le K \sum_{i = 1}^\infty \frac{\gamma_i^\lambda}{i}
    < \infty \\
    & 
    ~~ \mbox{so} ~~
    \sum_{i = 1}^\infty \frac{1}{i}
    \epsilon_i^2 \indic{\tau_R > n} < \infty
    ~ \mbox{a.s.}
  \end{align*}
by Condition~\eqref{eqn:lazy-gamma}. 
%  
%  Since
%  \begin{align*}
%\left\{ \sup_i |\Delta_i| \leq R \right\} \cap \left\{ \sum_{i=1}^\infty \frac{|\Delta_i|^{1+\lambda} \indic{\tau_T > n} }{\sqrt{i}} < \infty  \right\}  
%\subset \left\{ \sum_{i=1}^\infty \frac{|\Delta_i|^{1+\lambda}}{\sqrt{i}} < \infty  \right\}
%  \end{align*}
%it follows from \eqref{eq:sup_delta} and by the arbitrary choice of $\epsilon$ that
  %\newtext{As it follows from \cite[Eq.~(A11)]{polyak1992acceleration} that $\Prob\left( \sup |\Delta_t| < \infty  \right)=1$}, we obtain
%As Lemma~\ref{lemma:converging-power-sum} implies in particular that $\{\Delta_i\}$ is a.s. bounded, 
 As in the proof of Theorems~2 and~4 in \cite{polyak1992acceleration}, the Robbins-Siegmund Theorem \cite{robbins1971convergence} applied to the increment of $|\Delta_t|^2$ implies that for every 
  $\epsilon>0$ there exists some $R'>0$ such that
  \begin{align}
  \label{eq:sup_delta}
  \Prob\left(\sup_i |\Delta_i| \le R' \right) \ge 1-\epsilon.
  \end{align} 
Consequently, there exists some $R'' < \infty$ such that $\tau_{R''} = \infty$. 
We obtain that 
 \begin{equation*}
    \sum_{i = 1}^\infty \frac{1}{i}
    \varepsilon_i^2  < \infty
    ~~ \mbox{a.s.}.
  \end{equation*}
  Applying the square integrable martingale convergence
  theorem of \cite[Ex.~5.3.35]{Dembo16}, we have
  \begin{equation*}
    \frac{1}{\sqrt{n}} \sum_{i = 1}^n
    \left(\frac{1}{\psi'(0)} + w_i^n\right) \epsilon_i \cas 0,
  \end{equation*}
  so that both equations~\eqref{eq:PJ_proof1} and~\eqref{eq:PJ_proof2}
  converge almost surely to 0.
\end{proof}
